### 2022-11-25 MSDS 6371 Project - Yule and Kondakindi.qmd
Initial commit

---
title: "MSDS 6371 Project - Fall 2022"
author: "Andrew Yule, Krithika Kondakindi"
format: pdf
editor: visual
editor_options: 
  chunk_output_type: console
---

# Introduction

# Data Description

# Analysis

### Question 1

Brief introduction to the questions of interest and the setting of the problem.

#### Restatement of the Problem

(Where did the data come from? How big is it? How many observations? Where can we find out more? What are the specific variables that we need to know with respect to your analysis?)

#### Build and Fit the Model

#### Checking the Assumptions

Residual Plots Influential point analysis (Cook's D and Leverage) Make sure to address each assumption.

#### Comparing Competing Models

Adj R2 Internal CV Press

#### Parameters

Estimates Interpretation Confidence Intervals

#### Conclusion

A short summary of the analysis.

### R Shiny: Price v. Living Area Chart

### Question 2

#### Restatement of Problem

#### Model Selection

Type of Selection Stepwise Forward Backward Custom

#### Checking Assumptions

Residual Plots Influential point analysis (Cook's D and Leverage) Make sure to address each assumption

#### Comparing Competing Models

Adj R2 Internal CV Press\
Kaggle Score

#### Conclusion

A short summary of the analysis

# Appendix

```{r}
library(tidyverse)
```

Read training and test data and perform any cleaning steps
```{r}
housesTrain = read_csv("train.csv")
housesTest = read_csv("test.csv")

remove(houses)
```