### 2022-12-01 Analysis 2.qmd
Analysis 2: Moved code from main markdown file into separate one to focus on Analysis 2 question

---
title: "Analysis 2"
author: "Andrew Yule"
format: pdf
editor: visual
editor_options: 
  chunk_output_type: console
---

Load required libraries

```{r}
library(tidyverse)
library(broom)
library(terra)
```

Read training and test data and perform any cleaning steps

```{r}
housesTrain = read_csv("train.csv")
housesTest = read_csv("test.csv")

# Remove ID column only from the training set
housesTrain = select(housesTrain, -Id)

# Find which columns have missing values between the training and test data sets
colsToDrop = housesTrain |>
  select_if(function(x) any(is.na(x))) |>
  colnames()
colsToDrop2 = housesTest |>
  select_if(function(x) any(is.na(x))) |>
  colnames()

# Remove any of the columns that had missing values
housesTrain =  housesTrain |>
  select(-starts_with(c(colsToDrop, colsToDrop2)))
housesTest =  housesTest |>
  select(-starts_with(c(colsToDrop, colsToDrop2)))

hist(housesTrain$LotArea)

# From the plot below, LotArea should likely cut any value off at 30,000 sq ft
housesTrain$LotArea = clamp(housesTrain$LotArea, lower = 0, upper = 30000)
housesTest$LotArea = clamp(housesTest$LotArea, lower = 0, upper = 30000)

# From the plot below, GrLivArea should likely cut any value off at 4000 sq ft
housesTrain$GrLivArea = clamp(housesTrain$GrLivArea, lower = 0, upper = 4000)
housesTest$GrLivArea = clamp(housesTest$GrLivArea, lower = 0, upper = 4000)

# From the plot below, MSSubClass looks to have a quadratic relationship

# From the plot below, LotArea should should have a log relationship, which means each value should be greater than or equal to 1
housesTrain$LotArea = clamp(housesTrain$LotArea, lower = 1)
housesTest$LotArea = clamp(housesTest$LotArea, lower = 1)

# From the plot below, OpenPorchSF should should have a log relationship, which means each value should be greater than or equal to 1
#housesTrain$OpenPorchSF = clamp(housesTrain$OpenPorchSF, lower = 1)
#housesTest$OpenPorchSF = clamp(housesTest$OpenPorchSF, lower = 1)

# From the plot below, MiscVal should should have a log relationship, which means each value should be greater than or equal to 1
housesTrain$MiscVal = clamp(housesTrain$MiscVal, lower = 1)
housesTest$MiscVal = clamp(housesTest$MiscVal, lower = 1)

# From the plot below, PoolArea should should have a log relationship, which means each value should be greater than or equal to 1
#housesTrain$PoolArea = clamp(housesTrain$PoolArea, lower = 1)
#housesTest$PoolArea = clamp(housesTest$PoolArea, lower = 1)

# Screen for outliers by looking at the histograms of any variables that are all numeric
housesTrain |> 
  select(where(is.numeric)) |>
  gather() |>
  ggplot(aes(x = value)) + 
    facet_wrap(~key, scales = "free") + 
    geom_histogram()

# MoSold should be a factor instead of integer
housesTrain$MoSold = factor(housesTrain$MoSold, levels = seq(1, 12))
housesTest$MoSold = factor(housesTest$MoSold, levels = seq(1, 12))

# The sale price has a wide variance. Applying a log transformation helps
housesTrain |>
  ggplot(aes(x = log(SalePrice))) +
  geom_histogram()

# Log of SalePrice should be used to reduce the variance
housesTrain$SalePrice = log(housesTrain$SalePrice)

```

Full model creation

```{r}
housesTrain |> count(`3SsnPorch`)

fit = lm(SalePrice ~ I(MSSubClass^2) + log(LotArea) + Street + LotShape + LandContour + LotConfig + LandSlope + Neighborhood + Condition1 + Condition2 + BldgType + HouseStyle + OverallQual + OverallCond + YearBuilt + YearRemodAdd + RoofStyle + RoofMatl + ExterQual + ExterCond + Foundation + Heating + HeatingQC + CentralAir + `1stFlrSF` + `2ndFlrSF` + LowQualFinSF + GrLivArea + FullBath + HalfBath + BedroomAbvGr + KitchenAbvGr + TotRmsAbvGrd + Fireplaces + PavedDrive + WoodDeckSF + OpenPorchSF + EnclosedPorch + `3SsnPorch` + ScreenPorch + PoolArea + log(MiscVal) + MoSold + YrSold + SaleCondition, data = housesTrain)
summary(fit)

```

Test the basic model against the test data set and submit to Kaggle to see the score

```{r}
results = housesTest
results$SalePrice = exp(predict.lm(fit, newdata = housesTest))
results = results |>
  select(Id, SalePrice)

# Export
write_csv(results, "test_submission_10.csv")
```

Try using forward stepping approach

```{r}
# Define intercept-only model, this is also just the median sale price
intercept_only = lm(SalePrice ~ 1, data = housesTrain)

# Define model with all predictors
all = lm(SalePrice ~ ., data = housesTrain)

# Perform forward step-wise regression
forward = step(intercept_only, direction = 'forward', scope = formula(all), trace = 0)

# View final model
summary(forward)

```

Test the forward model against the test data set and submit to Kaggle

```{r}
results = housesTest
results$SalePrice = exp(predict.lm(forward, newdata = housesTest))
results = results |>
  select(Id, SalePrice)

# Export
write_csv(results, "test_submission_forward_1.csv")
```

Try using backward stepping approach

```{r}
# Define model with all predictors
all = lm(SalePrice ~ ., data = housesTrain)

# Perform forward step-wise regression
backward = step(all, direction = 'backward', scope = formula(all), trace = 0)

# View final model
summary(backward)

```

Test the backward model against the test data set and submit to Kaggle

```{r}
results = housesTest
results$SalePrice = exp(predict.lm(backward, newdata = housesTest))
results = results |>
  select(Id, SalePrice)

# Export
write_csv(results, "test_submission_backward_1.csv")
```

Try using stepwise stepping approach

```{r}
# Define intercept-only model, this is also just the average fat amount
intercept_only = lm(SalePrice ~ 1, data = housesTrain)

# Define model with all predictors
all = lm(SalePrice ~ ., data = housesTrain)

# Perform forward step-wise regression
both = step(intercept_only, direction = 'both', scope = formula(all), trace = 0)

# View final model
summary(both)
```

Test the stepwise model against the test data set and submit to Kaggle

```{r}
results = housesTest
results$SalePrice = exp(predict.lm(both, newdata = housesTest))
results = results |>
  select(Id, SalePrice)

# Export
write_csv(results, "test_submission_stepwise_1.csv")
```