### 2022-12-12 Analysis 1.qmd
Updates

---
title: "Analysis 1"
format: pdf
editor: visual
editor_options: 
  chunk_output_type: console
---

Define a function to determine the PRESS of each model
```{r}
PRESS = function(model) {
    i = residuals(model)/(1 - lm.influence(model)$hat)
    sum(i^2)
}
```

Create models for Sales Price for the neighborhood Brookside, Edwards, and North Ames neighborhoods.
```{r}

housesTrain |>
  ggplot(aes(x = GrLivArea, y = SalePrice)) +
  geom_point()

housesTrainNeighborhood = housesTrain |>
  filter(Neighborhood %in% c("BrkSide", "Edwards", "NAmes"))

housesTrainNeighborhood$Neighborhood = factor(housesTrainNeighborhood$Neighborhood, levels = c("NAmes", "BrkSide", "Edwards"))

housesTrainNeighborhood |>
  group_by(Neighborhood) |>
  summarize(n = n())

housesTrainNeighborhood |>
  ggplot(aes(x = GrLivArea, y = SalePrice, color = Neighborhood)) +
  geom_point() +
  facet_wrap(~Neighborhood) +
  theme(legend.position="none") +
  xlab("Log Value of Living Area Square Footage") +
  ylab("Log Value of Sales Price") +
  ggtitle("Home Sales Price and Living Area")

model1 = lm(SalePrice ~ Neighborhood, data = housesTrainNeighborhood)
summary(model1)
PRESS(model1)

model2 = lm(SalePrice ~ Neighborhood + GrLivArea, data = housesTrainNeighborhood)
summary(model2)
PRESS(model2)

model3 = lm(SalePrice ~ Neighborhood + GrLivArea + Neighborhood:GrLivArea, data = housesTrainNeighborhood)
summary(model3)
PRESS(model3)

plot(model3)

housesTrainNeighborhood |>
  select(Neighborhood, GrLivArea, SalePrice) |>
  mutate(SalePricePredicted = predict(model3, newdata = tibble(Neighborhood = Neighborhood, GrLivArea = GrLivArea))) |>
  ggplot() +
  geom_point(aes(x = exp(GrLivArea), y = exp(SalePrice), color = Neighborhood)) +
  geom_line(aes(x = exp(GrLivArea), y = exp(SalePricePredicted), color = Neighborhood)) +
  facet_wrap(~Neighborhood) +
  theme(legend.position="none")

```